---
title: "Project"
author: "Sharon and Catherine"
date: '2022-04-14'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

This project seeks to understand how terrorism has impacted the world.
The following questions will be used to understand the dataset;
1. Has global terrorism reduced or increased over the years?(line chart)
2. How do terrorism fatalities compare across countries in the world? (Choropleth map)
3. Which region has highly been affected by terrorism?
4. What’s the happiness index and GDP of the country highly affected by terrorism? Compare

## Load Packages and Import Data
The chunk below contains the packages that will be needed for this project.

```{r Load Packages}
#Load necessary packages
library(tidyverse) #For performing data transformation
library(ggplot2) #For creating charts
library(readr) #For importing data
library(dplyr) #For performing data transformation
```

```{r Import Data}
#Import dataset containing terrorism fatalities
terrorism_fatalities <- read_csv("fatalities-from-terrorism.csv")
```

```{r Display Dataset}
#Display the first five rows of the imported dataset
head(terrorism_fatalities)
```

##Data Transformation

Check for the number of rows and columns
```{r Number of rows}
nrow(terrorism_fatalities)
```

```{r Number of columns}
ncol(terrorism_fatalities)
```


```{r Missing Values}
#Check for missing values
sum(is.na(terrorism_fatalities))
```

The terrorism fatalities dataset has 153 missing values. Given that the dataset has 3814 entries, 







