---
title: "Analysis of Terrorism Fatalities Across the World"
author: "Sharon and Catherine"
date: '2022-04-14'
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



This project seeks to understand how terrorism has impacted the world.\ 

To accomplish this project, we seek to address the following questions:\
1. Which regions have been most/least affected by terrorism? Has terrorism, in the top 5 most affected countries, reduced or increased over the years (2000-2017)?\
2. What is the impact of terrorism on countries most/least affected by terrorism? Is the impact positive or negative?\
3. How do terrorism fatalities compare across the world?\
4. What is the most prevalent type of terrorism?
5. Which sectors and nationalities are highly targeted by terrorism activities?\

# Exploratory Data Analysis (EDA) #

## Load Packages and Import Data.

```{r Load Packages, message=FALSE}
#Load necessary packages
library(tidyverse) 
library(ggplot2) 
library(readr) 
library(dplyr) 
```

```{r Import Data, message = FALSE}
#Import dataset containing terrorism fatalities
terrorism_fatalities <- readr::read_csv("fatalities-from-terrorism.csv", show_col_types = FALSE)
```

```{r Display Dataset}
#Display the first five rows of the imported dataset
as_tibble(terrorism_fatalities)
head(terrorism_fatalities)
```


Check the number of rows and columns.\

```{r Rows}
#Number of rows
nrow(terrorism_fatalities)
```

```{r Columns}
#Number of columns
ncol(terrorism_fatalities)
```


Check if there are any missing values.\

```{r Missing Values}
#Check for missing values
#sum(is.na(terrorism_fatalities))

colSums(is.na(terrorism_fatalities))
```

The terrorism fatalities dataset has no missing values.\

Check for unique values in the 'Entity' column.\

```{r}
#Unique values in the dataset
unique(terrorism_fatalities[c("Entity")])
```

Rename 'Entity' column to 'Country'.\

```{r Rename}
terrorism_fatalities <- terrorism_fatalities %>%
  rename(Country = Entity)
```

Summary statistics of the dataset.\

```{r Summary}
summary_stat <- terrorism_fatalities %>% 
  group_by(Country) %>%
  summarise(
    count = n(), #Number of rows 
    avg = mean(Fatalities), 
    minimum = min(Fatalities),
    maximum = max(Fatalities)
  ) %>% 
  arrange()
summary_stat

```



# Data Transformation #

### 1. Which regions have been most/least affected by terrorism? Has terrorism, in the top 5 most affected countries, reduced or increased over the years (2000-2017)?\

Reorder the dataset to check which countries have had the highest number of fatalities as a result of terrorism.\

```{r Arrange}
highest <- terrorism_fatalities %>% 
  group_by(Country) %>%
  summarise(maximum = max(Fatalities)) %>% 
  arrange(desc(maximum)) #Ordering the values from the highest to the lowest

highest
```

**Above, we can see that Iraq has had the highest number of fatalities resulting from terrorism, followed by Nigeria, Afghanistan, Algeria and Syria.**\

The regions that have been least affected by terrorism.\

```{r}
lowest <- terrorism_fatalities %>% 
  group_by(Country) %>%
  summarise(minimum = max(Fatalities)) %>% 
  arrange(minimum) 

lowest
```

**The five countries that have been least affected by terrorism are Antigua & Barbuda, Benin, Brunei, Falkland Islands and French Polynesia.**\

The top five countries that have been most affected by terrorism.\

```{r}
highest_terrorism <- highest %>%
filter(maximum >= 3924)

highest_terrorism
```

**The top 5 most affected countries by terrorism are Iraq, Nigeria, Afghanistan, Algeria and Syria.**\

```{r Merge dataset}
terrorism_high <- highest_terrorism %>%
  left_join(terrorism_fatalities, by = "Country")

terrorism_high
```

Visualize the trend in terrorism fatalities of the top 5 most affected countries between Years 2000 and 2017.\

```{r Geom_line, echo=FALSE}
plot_data <- terrorism_high %>% 
  filter(Year >= 2000)

ggplot(terrorism_high, aes(x = Year, y = Fatalities, group=Country, color=Country)) + 
  geom_line()+
  ggtitle("Terrorism Fatalities of the Top Five Most Affected Countries 
          Between (2000-2017)") +
  xlab("Year") + 
  ylab("Number of Fatalities")

```

**From the chart above, we can see that the number of fatalities as a result of terrorism were on the rise over the years for the top five most affected countries, and they started declining towards 2017.**\ 

### 2. What is the impact of terrorism on countries most/least affected by terrorism? Is the impact positive or negative?

To determine the impact of terrorism, we compared the life satisfaction and GDP Per Capita of the countries most affected by terrorism with the countries least affected by terrorism.\

Load the dataset containing the GDP Per Capita and Life Satisfaction of each country.\

```{r}
#Import dataset containing life satisfaction and gdp of countries
gdp_happiness <- readr::read_csv("gdp-vs-life-satisfaction.csv", show_col_types = FALSE)

gdp_happiness 
```

Join 'gdp_happiness' dataset with 'terrorism_fatalities' dataset.\

```{r}
merged_data <- gdp_happiness %>% 
  full_join(terrorism_fatalities, by = "Country")

merged_data
```

Compute the average life satisfaction, fatalities and GDP of each country.\

```{r Average}
average_data <- merged_data %>% 
  group_by(Country) %>% 
  summarise(
    average_satisfaction = mean(Life_Satisfaction, na.rm=TRUE),
    average_gdp = mean(GDP, na.rm=TRUE),
    average_fatalities = mean(Fatalities, na.rm=TRUE))

average_data
```

The countries with the highest life satisfaction\

```{r}
high_to_low <- average_data %>% 
  select(Country, average_satisfaction) %>%
  arrange(desc(average_satisfaction)) #Ordering the values from the highest to the lowest

high_to_low
```

**Denmark has had the highest average life satisfaction then Norway and Switzerland. The countries most affected by terrorism do not appear in the list of countries with high average life satisfaction.**\

Filter top five most affected and least affected countries by terrorism? 

```{r}
countries_ranked <- highest %>%
  select(Country, maximum) %>% 
  filter(Country %in% c("Iraq", "Nigeria", "Afghanistan", "Algeria", "Syria", 
                        "Benin", "Iceland", "Mauritius", "Brunei", "Seychelles"))

countries_ranked
```

```{r}
joined_data <- countries_ranked %>% 
  left_join(merged_data, by = "Country")

joined_data
```


```{r}
comparison_data <- joined_data %>%
  group_by(Country) %>% 
  summarise(
    avg_satisfaction = mean(Life_Satisfaction, na.rm = TRUE),
    avg_fatalities = mean(Fatalities, na.rm = TRUE),
    avg_gdp = mean(GDP, na.rm = TRUE))
comparison_data
```
The country with the highest average life satisfaction, Denmark, had a value of 7.695750. On the other hand, the top 5 countries that have had high terrorism fatalities, have average life satisfaction values ranging from 5.555167-3.806700.This implies that the aforesaid countries have low life satisfaction.\

```{r Geom_bar}
#import library gridExtra to help in joining the plots
library(gridExtra)

#fatalities_plot <- comparison_data %>% 
  #ggplot(mapping = aes(x = avg_fatalities, y = Country))+
  #geom_bar(stat="identity", width=1, color="white",
         #fill=rgb(0.8,0.4,0.1,0.7))+
  #xlab("Number of Fatalities")

satisfaction_plot <- comparison_data %>% 
  ggplot(mapping = aes(x = avg_satisfaction, y = Country))+
  geom_bar(stat="identity", width=1, color="white",
         fill=rgb(0.1,0.8,0.1,0.7))+
  xlab("Life Satisfaction")
 
gdp_plot <- comparison_data %>% 
  ggplot(mapping = aes(x = avg_gdp, y = Country))+
  geom_bar(stat="identity", width=1, color="white",
         fill=rgb(0.1,0.4,0.5,0.7))+
  xlab("GDP Per Capita")
 
# Use grid.arrange to put plots in columns
grid.arrange(gdp_plot,satisfaction_plot, nrow = 2, 
             top = "Comparison of Life Satisfaction and GDP of Countries 
             Most/Least Affected by Terrorism")
```

We can tell from the chart above that the countries that were **least affected by terrorism (Iceland, Mauritius, Brunei and Seychelles) have high GDP per capita** compared to the countries that were most affected by terrorism. Afghanistan, Iraq and Nigeria have low GDP per capita. More so, countries **least affected by terrorism (Iceland and Mauritius) have higher life satisfaction** compared  to countries most affected by terrorism. Therefore, from the outcome above, we can infer that presence of terrorism activities has a negative impact on countries. 

### 3. How do terrorism fatalities compare across the world?

Compute the average terrorism rate for each country.\

```{r}
#Group by Country and find the terrorism rate for each country
terrorism_fatalities <- terrorism_fatalities %>% 
  group_by(Country) %>% 
  summarise(Fatalities=mean(Fatalities))
```

Rename 'Country' column to 'region'.\

```{r}
terrorism_fatalities <- terrorism_fatalities %>%
  rename(region = Country)
```

Plot a choropleth map to show how terrorism compare across countries.\

```{r Choropleth Map}
#Convert from tibble to dataframe
terrorism_fatalities <- as.data.frame(terrorism_fatalities)

#Data containing longitudes and latitudes to be used in mapping the data
global_map <- map_data("world")

#Merge the "world" map_data to the terrorism_fatalities dataset
merged_data <- left_join(global_map, terrorism_fatalities, by = "region")

#Plot the map
ggplot(merged_data, aes(long, lat, label = region ,group = group))+
  geom_polygon(aes(fill = Fatalities), color = "white")+
  ggtitle("Fatalities Resulting From Terrorism Activities Across the World")+
  scale_fill_viridis_c(option = "C")+
  theme(#panel.background = element_rect(fill = "white"),
        #plot.background = element_rect(fill = "white"),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        #rect = element_blank(),
        #panel.border =  element_blank(),
        legend.title= element_text(
          hjust = 0.4 ,vjust=0.3, size=10,family = "Times"),
        legend.text = element_text(
          hjust = 0.4 ,vjust=2, size=8,family = "Times"))
  

```

From the map above, we can compare the fatalities as a result terrorism across the globe. We can see that some regions in Africa and Asia have highly been affected by terrorism.

### 4. What is the most prevalent type of terrorism?

```{r}
#Import dataset containing the types of terrorism
types_of_terrorism <- readr::read_csv("types-of-terrorism.csv", show_col_types = FALSE)

types_of_terrorism 
```

```{r}
#Unique values in the dataset
unique(types_of_terrorism[c("terrorism_type")])
```

There are nine major types of terrorism as shown above.


```{r}
rank_terrorism <- types_of_terrorism %>%
  group_by(terrorism_type) %>% 
  summarise(rank_terrorism = n())%>% 
  arrange(desc(rank_terrorism))

rank_terrorism
```

** Bombing/Explosion has been the most prevalent type of terrorism

### 5. Which sectors and nationalities are highly targeted by terrorism activities?\

```{r}
rank_sector <- types_of_terrorism %>%
  group_by(targeted_sectors) %>% 
  summarise(rank_sector = n())%>% 
  arrange(desc(rank_sector))

rank_sector
```

The top five sectors that have been a target of terrorism activities are private citizens, military, police, government and business respectively.

```{r}
rank_nationalities <- types_of_terrorism %>%
  group_by(targeted_nationalities) %>% 
  summarise(rank_nationalities = n())%>% 
  arrange(desc(rank_nationalities))

rank_nationalities
```

The top five nationalities that have been a target of terrorism activities are Iraq, Pakistan, Afghanistan, India and Columbia.

# Conclusion #

From the analysis above, it is evident that Asian countries (i.e Iraq, Afghanistan, Syria and Pakistan) have mostly been affected by terrorism. More so, we can infer that the presence of terrorism activities has had a negative impact, not only on the countries in general (i.e the GDP Per Capita of the countries), but also on the wellbeing of the citizens (i.e Life Satisfaction) belonging to those countries.




